---
title: "01-vectors-atomic"
output: html_notebook
---
This is the first video in my YouTube playlist rstats: introduction.

Vectors (come in two flavors:)
 * Atomic: logical, integer, double, character (these are types or models)
 * List (Data Frame or tibble)
 
An ATOMIC vector elements of only one type; i.e., either logical, integer, double or character. A list can contain a mix of types, including the element in a list can be another list (aka, recursive).

```{r}
three_dice <- c(2, 3, 6) # creates a vector with length of 3
three_dice
typeof(three_dice) # double
length(three_dice)

three_dice[2] # retreives the 2nd elements in the vector
str(three_dice) # STRucture is very useful

three_more <- c(1, 5, 2)
three_dice + three_more # adding two vectors element-wise
three_dice <- c(2L, 3L, 6L) # specifies type = integer
str(three_dice)

one_die <- 4 # a vector of length = 1 is called a scalar
str(one_die)
three_dice + one_die # the shorter vector is RECYCLED
three_dice

one_die <- "four" # character
two_dice <- c("two", "four")
str(two_dice)
length(two_dice)
two_dice[2]

coin <- c(FALSE, FALSE, TRUE, FALSE) # logical vector
str(coin)
sum(coin); mean(coin) # FALSE = 0 and TRUE = 1

```

Coercion and Testing
```{r}
# Coercion: least to most flexible: 
#                   logical -> integer -> double -> character

coin <- c(FALSE, TRUE, FALSE, FALSE)
coin <- c(FALSE, 1, TRUE, 0)
str(coin)
coin <- c(FALSE, 1, "heads", "tails")
str(coin)
coin[3]

# Testing
is.numeric(coin)
is.character(coin)
is.numeric(three_dice)

```

