---
title: "02-vectors-list"
output: html_notebook
---

This script is used my my video: R Programming: Introduction: List data structure (R Intro-02).
The video is located at https://youtu.be/_EhiIBKMJi4 

Vectors (1-dimensional data structures)

 * Atomic: elements all SAME type (logical, integer, double, character)  * List: elements of ANY type

```{r}
li <- list (
  c("AAPL", "BAC", "FB", "SBUX"),
  1:10,
  c(TRUE, TRUE, TRUE, FALSE),
  list(1:3, c("a", "b", "c")),
  c(mean, sum)
)

# li[n] returns a list
# li[[n]] returns element

li[1]
li[[1]]


```

Assign NAMES to elements in the list (names is a char vector ATTRIBUTE)

```{r}
nums <- 1:9
li <- list (
  stocks = c("AAPL", "BAC", "FB", "SBUX"),
  nums = nums,
  own = c(TRUE, TRUE, TRUE, FALSE),
  sub = list(1:3, c("a", "b", "c")),
  funcs = c(avg = mean, sum)
)
names(li)
li[c("stocks","own")]

li[["stocks"]]

li$stocks
li$own

li$funcs$avg(1:10)

```

Name-value association (per R Cookbook); aka, lookup tables
```{r}
ticker <- list(AAPL = "Apple", BAC = "Bank of America", FB = "Facebook", SBUX = "Starbucks")
cat("The company is called", ticker$FB)

```

Regression model returns a LIST
```{r}
women
model <-  lm(formula = weight ~ height, data = women)
summary(model)
str(model)
model[["coefficients"]]
model$coefficients
model$coefficients[2]
model$coefficients["height"]


```

