---
title: "03-data-frames"
output: html_notebook
---

This script is used my my video: R Programming: Introduction: Data frames (R Intro-03).
The video is located at https://youtu.be/oVhDN7TfC08

A data frame is a list of equal-length vectors.

The video has three parts

* Create a data frame
* Basic subseting (selecting) from a data frame
* Tibble is a convenient improvement on the data frame (its class is both dataframe and tibble). Tibble is part of the tidyverse

Create a data frame 

```{r echo=TRUE}
# character vector
company <- c("Apple","Blackrock","CVS Health", "Enbridge", "Lennar", "Starbucks")

# create a date frame
stocks <- data.frame(
  company = company,
  sector = c("tech", "financial","healthcare","energy", "cyclical", "cyclical"),
  revenue = c(261.6, 14.2, 194.6, 34.5, 20.6, 25.3),
  buy = c(TRUE, TRUE, FALSE, TRUE, TRUE, FALSE), 
  stringsAsFactors = FALSE 
)

# easiest way to add column to df, but must be length = 6 observations
pe_ratio = c(14.3, 15.7, 7.5, 19.1, 8.0, 24.8)
stocks$pe <- pe_ratio

# can also cbind() or rbind()
ticker <- c("AAPL", "BLK", "CVS", "ENB", "LEN", "SBUX")
stocks <- cbind(ticker, stocks, stringsAsFactors = FALSE)
yext <- list("YEXT", "Yext Inc", "tech", 0.23, TRUE, NA)
stocks <- rbind(stocks, yext, stringsAsFactors = FALSE)

# list-columns
li <- list(1, 1:2, 1:3, 1:4, 1:5, 1:6, 1:7)
stocks$li <- li

# stocks <- stocks[-1]
stocks$ticker <- NULL

colnames(stocks); names(stocks)
rownames(stocks)
nrow(stocks); ncol(stocks) # matrix-like

rownames(stocks) <- ticker # but actually undesirable see 
ticker <- append(ticker, "YEXT")

stocks$ticker <- rownames(stocks) # put rows back into df
rownames(stocks) <- NULL
```

Subset a data frame
```{r echo=TRUE}
eg <- stocks[1] # data frame

eg <- stocks[[1]] # vector
stocks$company

# matrix style
# stocks[rows, columns]
eg <- stocks[2:4, 1:2]
eg <- stocks[, 1:2]

eg <- subset(stocks, select = c(sector, revenue), subset = revenue > 200)
```

Tibble https://cran.r-project.org/web/packages/tibble/vignettes/tibble.html

* Does not change inputs. Not needed: stringsAsFactors = FALSE
* Create list columns inside tibble (as above)
* Lazy arguments (below)
* Better printout and etc (strict subsetting)

```{r echo=TRUE}
library(tibble)
stocks_tbl <- as.tibble(stocks)
str(stocks_tbl)

stocks_tbl <- tibble(
  company = c("Apple","Blackrock","CVS Health", "Enbridge", "Lennar", "Starbucks"),
  sector = c("tech", "financial","healthcare","energy", "cyclical", "cyclical"),
  revenue = c(261.6, 14.2, 194.6, 34.5, 20.6, 25.3),
  buy = c(TRUE, TRUE, FALSE, TRUE, TRUE, FALSE),
  li = list(1:2, 1:3, 1:4, 1:5, 1:6, 1:7)
)
stocks_tbl

eg <- tibble(x = 1:5, y = x^2)

cars <- as.tibble(mtcars)
```

